### 2016-11-09 Manuscript/SW_early_life_telomeres.Rmd
eds

###Results
We measured telomere lengths using a total of `r nrow(dd)` unique samples from juvenile and adult Seychelles warblers. Efficiencies (mean $\pm$ s.d.) for our telomere and GAPDH reactions were (`r round2(mean(dd$EffTelomere))` $\pm$ `r round2(sd(dd$EffTelomere))`) and (`r round2(mean(dd$EffGAPDH))` $\pm$ `r round2(sd(dd$EffGAPDH))`) respectively. Inter-plate repeatability of RTL, based on `r nrow(subset(ddL,Group == 'Within samples'))` samples measured at least twice, was 0.78 (CI = 0.73-0.82).

#####Telomere dynamics and age
Using the cross-sectional data RTL decreased with age across the entire lifespan of the Seychelles warbler (LMM: t = `r round2(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,5])`). We found that modelling a non-linear effect age explained almost double the variation in telomere length (R^2^ of linear regression of telomere length and log age = `r round3(summary(lm(RTL~log(Agemonths),data=dd))$r.squared)` than a linear model `r round3(summary(lm(RTL~Agemonths,data=dd))$r.squared)`. Also worth noting is that while the overall trend was for decreasing telomere length, there were clear increases in cross-sectional telomere length after one year of age and at several points later in life (Fig. 1A).

Longitudinal data showed that early-life RTL was weakly, but significantly, related to adult RTL (R^2^ = `r round3(summary(lm(RTL1~RTL,data = dd3))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,3])`; P = `r round3(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,4])`; Fig. 1B). As with cross-sectional RTL, $\Delta$RTL decreased with age (R^2^ = `r round3(summary(lm(DeltaRTL~Agemonths,data = dd3))$r.squared)`; t = `r round2(summary(lm(DeltaRTL~Agemonths,data = dd3))$coefficients[2,3])`; P = `r round3(summary(lm(DeltaRTL~Agemonths,data = dd3))$coefficients[2,4])`), suggesting that telomere shortening was highest shortly after the nestling phase, after which both telomere shortening and telomere lengthening were observed.

To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we compared variance in telomere length among repeat measurements of samples to the variance observed among different samples of the same individual (see methods for details). We found significantly higher variance in telomere length over individual lifetimes compared to among sample replicates (Levene's test: F = `r round2(leveneTest(DeltaRTL~Group,data=ddL)['F value'][1,1])`; P `r round3(leveneTest(DeltaRTL~Group,data=ddL)['Pr(>F)'][1,1])`; Fig. 1C). Importantly, splitting the longitudinal data into instances of shortening ($\Delta$RTL < 0) and lenghtning ($\Delta$RTL > 0) revealed that not only did we observe significantly more shortening within individuals compared to within samples (Wilcoxon test: W = `r round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$statistic,0)`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$p.value)`), but also significantly more lengthening (W = `r as.character(round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$statistic,0))`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$p.value)`). We found no evidence that the likelihood of telomere lengthening was related to the follow up time between sampling events (GLM: t = `r round2(summary(glm(DeltaRTLF~TimeDiff,data=dd3,family = 'binomial'))$coefficients[2,3])`; P = `r round2(summary(glm(DeltaRTLF~TimeDiff,data=dd3,family = 'binomial'))$coefficients[2,4])`), and considering those samples that did show increases in $\Delta$RTL, the degree of increase was not related to follow up time (LM: t = `r round2(summary(lm(DeltaRTL~TimeDiff,data=subset(dd3,DeltaRTL>0)))$coefficients[2,3])`; P = `r round2(summary(lm(DeltaRTL~TimeDiff,data=subset(dd3,DeltaRTL>0)))$coefficients[2,4])`).

#####Early life environment and telomere dynamics
Early-life RTL varied significantly among cohorts (Likelihood ratio test: F = `r round2(anova(lme(RTL~factor(FieldPeriodID),data=juv,random=~1|BirdID))[2,3])`; P `r round3(anova(lme(RTL~Agemonths,data=juv,random=~1|BirdID))[2,4])`). Cohort-level variation in juvenile RTL was not related to adult population density (LMM controlling for age and cohort: t = `r round2(summary(lme(RTL~Agemonths+Density,data=subset(juv,!is.na(Density)),random=~1|FieldPeriodID))$tTable[3,4])`; P = `r round3(summary(lme(RTL~Agemonths+Density,data=subset(dd,!is.na(Density)),random=~1|BirdID))$tTable[3,5])`), or to temporal variation in  food availability (t = `r round2(summary(lme(RTL~Agemonths+Insect,data=dd,random=~1|BirdID/FieldPeriodID,na.action=na.exclude))$tTable[3,4])`; P = `r round3(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|BirdID,na.action=na.exclude))$tTable[3,5])`).

@@

Within cohorts, the top model explaining variation in early-life RTL contained age, territory quality, tarsus length, and the interaction between age and tarsus length (Table S1). Model-averaged confidence intervals of the age and tarsus x age effects did not overlap zero, but the tarsus and territory quality effects were not significant (Fig. 3A). Tarsus length was negatively related to RTL in nestlings, unrelated to RTL in fledglings, and positively related to RTL in subadults (Fig. 3B). Territory quality was negatively related to RTL, particularly in nestlings, although this effect was weak (Fig. 3C).

#####Telomere dynamics and survival
Early-life RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~RTL+LogAge+(1|PlateID)+(1|TerritoryID)+(1|LayYear),data=dd14,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL+LogAge+(1|TerritoryID)+(1|LayYear),data=dd14,family = 'binomial'))$coefficients[2,4])`), with no interaction effect between RTL and  age (estimate = `r round2(summary(glmer(SurvivedNext~RTL*LogAge+(1|TerritoryID)+(1|LayYear),data=dd14,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL*LogAge+(1|TerritoryID)+(1|LayYear),data=dd14,family = 'binomial'))$coefficients[4,4])`), suggesting that there were no age-specific effects of juvenile RTL on survival to adulthood.

summary(glmer(SurvivedNext~RTL+LogAge+Sex+Tarsus+(1|LayYear),data=subset(dd14,Agemonths>30),family = 'binomial'))

summary(coxph(Surv(RemainingLife,Died)~RTL+LogAge,data=dd8))

>>>>

###Results
We measured telomere lengths using a total of `r nrow(dd)` unique samples from juvenile and adult Seychelles warblers. Efficiencies (mean $\pm$ s.d.) for our telomere and GAPDH reactions were (`r round2(mean(dd$EffTelomere))` $\pm$ `r round2(sd(dd$EffTelomere))`) and (`r round2(mean(dd$EffGAPDH))` $\pm$ `r round2(sd(dd$EffGAPDH))`) respectively. Inter-plate repeatability of RTL, based on `r nrow(subset(ddL,Group == 'Within samples'))` samples measured at least twice, was `r round2(rptall$R)` (CI = `r round2(rptall$CI_emp[1])`-`r round2(rptall$CI_emp[2])`)

#####Telomere dynamics and age
Using the cross-sectional data, we found that RTL decreased with age in the Seychelles warbler (LMM: estimate $\pm$ s.e. = `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID/LayYear))$tTable[2,1])` $\pm$ `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID/LayYear))$tTable[2,2])`; t = `r round2(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID/LayYear))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID/LayYear))$tTable[2,5])`). We found that modelling a non-linear effect age explained almost double the variation in RTL (R^2^ of linear regression of telomere length and log age = `r round3(summary(lm(RTL~log(Agemonths),data=dd))$r.squared)` than a linear model `r round3(summary(lm(RTL~Agemonths,data=dd))$r.squared)`, suggesting that telomere length decreased at the greatest rate in early life. Accordingly, RTL decreased with age at a greater rate in juveniles (estimate $\pm$ s.e. = `r round3(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID/LayYear))$tTable[2,1],lessthan = F)` $\pm$ `r round3(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID/LayYear))$tTable[2,2])`) compared to adults (estimate $\pm$ s.e. = `r round3(summary(lme(RTL~Agemonths,data=subset(dd,Ageclass == 'A'),random=~1|BirdID/LayYear))$tTable[2,1],lessthan = F)` $\pm$ `r round3(summary(lme(RTL~Agemonths,data=subset(dd,Ageclass == 'A'),random=~1|BirdID/LayYear))$tTable[2,2])`). Also worth noting is that while the overall trend was for decreasing RTL, there were appoarent increases in cross-sectional telomere length after one year of age and at several points later in life (Fig. 1A).

Longitudinal data showed that early-life RTL was weakly, but significantly, related to adult RTL (R^2^ = `r round3(summary(lm(RTL1~RTL,data = dd3))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,3])`; P = `r round3(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,4])`; Fig. 1B). As with cross-sectional RTL, $\Delta$RTL decreased with age (R^2^ = `r round3(summary(lm(DeltaRTL~LogAge,data = dd3))$r.squared)`; t = `r round2(summary(lm(DeltaRTL~LogAge,data = dd3))$coefficients[2,3])`; P = `r round3(summary(lm(DeltaRTL~LogAge,data = dd3))$coefficients[2,4])`), suggesting that telomere shortening was highest in early life. Although both or cross-sectional and longitudinal data indicated a general trend of telomere shortening, we found that `r round((nrow(subset(dd3,DeltaRTL>0))/nrow(dd3))*100,0)% of our $\Delta$RTL measurments were positive, and thus indicative of telomere lengthening. To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we compared variance in telomere length among repeat measurements of samples to the variance observed among different samples of the same individual (see methods for details). We found significantly higher variance in telomere length over individual lifetimes compared to among sample replicates (Levene's test: F = `r round2(leveneTest(DeltaRTL~Group,data=ddL)['F value'][1,1])`; P `r round3(leveneTest(DeltaRTL~Group,data=ddL)['Pr(>F)'][1,1])`; Fig. 1C). Importantly, splitting the longitudinal data into instances of shortening ($\Delta$RTL < 0) and lenghtning ($\Delta$RTL > 0) revealed that not only did we observe significantly more shortening within individuals compared to within samples (Wilcoxon test: W = `r round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$statistic,0)`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$p.value)`), but also significantly more lengthening (W = `r as.character(round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$statistic,0))`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$p.value)`).

#####Spatiotemporal variation in telomere dynamics and the environment
There was significant variation in RTL among cohorts (), although we found no evidence that the relationshio between telomere length and age varied among cohorts. Rather, cohort-level differences in RTL were consistent over the lifespan of Seychelles warblers.

A linear mixed model of RTL in relation to social and ecological environmental variables revealed age, tarsus length, sex and annual variation in food availability were significantly related to RTL (Table 1). Model selection revealed that sex only appeared in models where tarsus length was also present (Table S1), and sex was no longer significant when tarsus length was removed from the full model (P = ). RTL increased with tarsus length in males, but decreased in females. In general, the model averaging approach yielded qualitatively identical results to the full LMM, with the same explanatory variables 'significant' in terms of being retained in top models, and having model-average confidence intervals not overlapping zero (Table S1; Fig. S1).

@@

### 2016-11-09 Manuscript/SW_early_life_telomeres_SOM.Rmd
eds

#Run the model
lm1 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID)+(1|TerritoryID)+(1|PlateID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

@@

nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'c.Season'] <- 'Season'
nams[nams == 'c.Season:z.Agemonths'] <- 'Age x season'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

```{r Table S2}
temp <- dd3

#Run the model
lm1 <- glmer(DeltaRTLF ~ Agemonths+(Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID) + (1|TerritoryID),
            data = temp,
            family = 'binomial',
            REML = FALSE)
lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

@@

```{r Figure S1}

temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)
temp$LayYear <- factor(temp$LayYear)
temp$TimeDiff <- temp$TimeDiff/365

FigS1 <- ggplot(temp,
       aes(x = TimeDiff,
           y = Loss,
           colour=`Age (months)`,
           fill = `Age (months)`)) +
  ylab(expression(Delta*'RTL')) + 
  xlab('Time (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
    geom_abline(slope = 0,intercept = 0,lty = 2)+
  annotate('text',x = 0,y = 1.75,label = 'B')

@@

>>>>

#Run the model
temp <- subset(dd,!is.na(Insect))

#Run the model
lm2 <- lmer(RTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|BirdID) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
t1 <- as.matrix(subset(lm3, delta<=6))

@@

nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

@@

```{r Table S2}
#Run the model
temp <- subset(dd3,!is.na(Insect))

#Run the model
lm2 <- lmer(DeltaRTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
t1 <- as.matrix(subset(lm3, delta<=6))

@@

```{r Figure S1}
temp <- subset(dd,!is.na(Insect))
temp$LogAge <- log(temp$Agemonths)

#Run the model
lm2 <- lmer(RTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|BirdID) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))
table2$myx <- factor(table2$myx,levels = table2$myx[order(table2$Estimate,decreasing = T)])

FigS1 <- forestplot(table2)+
  annotate('text',x = 0.6,y = 0.05,label = 'A')+
  theme(axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))

@@

```{r Figure S2}

#Run the model
temp <- subset(dd3,!is.na(Insect))

#Run the model
lm2 <- lmer(DeltaRTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))
table2$myx <- factor(table2$myx,levels = table2$myx[order(table2$Estimate,decreasing = T)])

FigS2 <- forestplot(table2)+
  annotate('text',x = 0.6,y = 0.05,label = 'A')+
  theme(axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))

FigS2

### 2016-11-09 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
eds

@@

Fig1A <- subset(dd,Age<15) %>%
  ddply(.(BirdID),
        mutate,
        nreps = length(RTL)) %>%
  subset(nreps>0) %>%
  subset(!duplicated(paste0(BirdID,Agemonths))) %>%
  transform(Agemonthsdec = Agemonths/12) %>%
  ggplot(aes(x = Agemonthsdec,y = RTL,group = BirdID,col=factor(BirdID)))+
  geom_point(alpha = 1,col='grey')+
geom_line(col = 'darkgrey',alpha = 0.2)+
  xlab('Age (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')+
  geom_smooth(col = 'black',fill = 'black',lwd = 0.5,aes(group = NULL),method = 'lm',formula = y~log(x))
  annotate('text',x = 0,y = 3,label = 'A')

@@

Fig2 <- ddply(dd,
               .(Agemonths,LayYear),
               summarise,
               meanRTL = mean(RTL),
               n = length(RTL))  %>%
  ddply(.(LayYear),
        mutate,
        nAges = length(meanRTL),
        minAge = min(Agemonths)) %>%
  subset(minAge == 1) %>%
  subset(n > 5) %>%
  transform(Cohort = factor(LayYear)) %>%
  ggplot(aes(x = Agemonths,y = meanRTL,col = Cohort))+
  xlab('Age (months)') +
  geom_point()+
  geom_line()+
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))
  annotate('text',x = 0,y = 1.4,label = 'B')

@@

temp <- subset(dd,!is.na(Insect))
temp$LogAge <- log(temp$Agemonths)

#Run the model
lm2 <- lmer(RTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|BirdID) + (1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.GroupSize'] <- 'Group Size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))
table2$myx <- factor(table2$myx,levels = table2$myx[order(table2$Estimate,decreasing = T)])

Fig3A <- forestplot(table2)+
  annotate('text',x = 0.6,y = 0.05,label = 'A')+
  theme(axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))

Fig3B <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           col = Sex)) +
  geom_point()+
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
  stat_smooth(method = 'lm')
  annotate('text',x = 17,y = 2.5,label = 'C')

Fig3C <- ggplot(temp,
       aes(
           y = RTL,
           x=Insect)) +
    ylab('RTL') + 
  xlab('Insect abundance') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
  geom_boxplot(aes(group = Insect),outlier.shape = NA)+
   stat_smooth(method = 'lm',col = 'darkgrey')
  annotate('text',x = 2.5,y = 2.5,label = 'D')

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
multiplot(Fig3A,Fig3B,Fig3C,layout = matrix(c(1,1,2,3),nrow = 2,byrow = T))
dev.off()

@@

```{r Figure 4}
dd14$r <- sample(1:nrow(dd14),replace = F)
temp <- dd14[order(dd14$BirdID,dd14$r,decreasing = F),] %>%
  subset(!duplicated(BirdID)) %>%
  subset(Died == 1)

Fig4A <- ggplot(temp,
       aes(x = Sex,
           col = factor(SurvivedNext),
           y = RTL)) +
  ylab('RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_boxplot(notch=T)
  annotate('text',x = 17,y = 2.5,label = 'C')

temp <- ddply(juv14,
              .(LayYear,Fledged),
              summarise,
              MeanTL = median(RTL),
              Survived = mean(SurvivedNext),
              n = length(RTL)) %>%
  subset(n > 9)

ggplot(temp,aes(x = MeanTL,y = Survived,col = Fledged))+
  geom_point()+
  stat_smooth(method = 'lm')

Fig4B <- ??

pdf('Figures/Figure 4.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
multiplot(Fig4A,Fig4B,layout = matrix(c(1,2),nrow = 2,byrow = T))
dev.off()

x <- subset(dd14,Fledged == 'Fledglings')

table(x$Sex,x$LayYear)

dd14$SexAge <- paste(dd14$Fledged,dd14$Sex)

ggplot(dd14,aes(x = Tarsus,y = RTL,col = SexAge))+
geom_smooth(method = 'lm')

>>>>

**Table1**
```{r Table 1}
temp <- subset(dd,!is.na(Insect))

#Run the model
lm2 <- lmer(RTL ~  LogAge + Tarsus + Helper + GroupSize + Sex + BodyMass + Insect + Density + (1|LayYear) + (1|BirdID) + (1|PlateID),
            data = temp,
            REML = FALSE)
CIs <- confint.merMod(lm2)
t1 <- data.frame(Estimate = fixef(lm2), s.e. = se.fixef(lm2), LCI = CIs[5:nrow(CIs),1], UCI = CIs[5:nrow(CIs),2])
kable(t1)

@@

Fig1A <- subset(dd,Age<15) %>%
  ddply(.(BirdID),
        mutate,
        nreps = length(RTL)) %>%
  subset(nreps>0) %>%
  subset(!duplicated(paste0(BirdID,Agemonths))) %>%
  ggplot(aes(x = Agemonths,y = RTL,group = BirdID,col=factor(BirdID)))+
  geom_point(alpha = 1,col='grey')+
geom_line(col = 'darkgrey',alpha = 0.2)+
  xlab('Age (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')+
  geom_smooth(col = 'black',fill = 'black',lwd = 0.5,aes(group = NULL),method = 'lm',formula = y~log(x))
  annotate('text',x = 0,y = 3,label = 'A')

@@

Fig2A <- transform(juv,Cohort = factor(LayYear)) %>%
  subset(Agemonths>(1/12)) %>%
  ggplot(aes(y = RTL,x = Cohort))+
  xlab('Cohort') +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))
  annotate('text',x = 0,y = 1.4,label = 'B')

Fig2B <- ddply(dd,
               .(Agemonths,LayYear),
               summarise,
               meanRTL = mean(RTL),
               n = length(RTL))  %>%
  ddply(.(LayYear),
        mutate,
        nAges = length(meanRTL),
        minAge = min(Agemonths)) %>%
  subset(LayYear %in% c(2001:2009)) %>%
  subset(n>4) %>%
  transform(Cohort = factor(LayYear)) %>%
  ggplot(aes(x = Agemonths,y = meanRTL,col = Cohort))+
  xlab('Age (years)') +
  ylab('Mean RTL') +
  geom_point()+
  geom_line()+
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')
  annotate('text',x = 0,y = 1.4,label = 'B')

@@

Fig3A <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           col = Sex)) +
  geom_point()+
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
  stat_smooth(method = 'lm')
  annotate('text',x = 17,y = 2.5,label = 'C')

Fig3B <- ggplot(temp,
       aes(
           y = RTL,
           x=Insect)) +
    ylab('RTL') + 
  xlab('Insect abundance') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
  geom_boxplot(aes(group = Insect),outlier.shape = NA)+
   stat_smooth(method = 'lm',col = 'darkgrey')
  annotate('text',x = 2.5,y = 2.5,label = 'D')

pdf('Figures/Figure 3.pdf',
    width = 5,
    height = 8,
    paper = 'a4')
multiplot(Fig3A,Fig3B)
dev.off()

@@
